Acceleration of Global Optimization Algorithm by Detecting Local Extrema Based on Machine Learning

### 2021-09-15 2021/Entropy/entropy.tex
rewrite some part of the Introduction

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
%\usepackage{inputenc}
%\usepackage[T2A,T1]{fontenc}
%\usepackage[english,russian]{babel}
%\usepackage{cmap}
%%%%

@@

There are several approaches to the construction of numerical methods for solving global optimization problems. 
%Надо ли добавить алгоритмы на основе метамоделей ?
Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters. Local optimization methods have a high convergence rate. At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions. 
Machine learning methods can be successfully applied to solve this problem.  
For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
In \cite{Cassioli2012} the area for starting the local method was allocated based on the classification of starting points using a support vector machine.

@@

Note that the algorithms of the last two classes do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set). Therefore, deterministic methods seem to be potentially more effective.

One of the efficient deterministic methods for solving multiextremal optimization problems is \textit{the information-statistical global search algorithm} \cite{Strongin2000}. This method initially proposed for solving unconstrained optimization problems was successfully generalized to the classes of optimization problems with non-convex constraints \cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}. For different versions of the algorithm, parallelization methods taking into account the architecture of modern computing systems were also suggested \cite{Barkalov2016,globalizerSystem,Strongin2018}.

@@

A typical situation for many applied problems is when a limited change in the vector of parameters $y$ causes a limited change in the values of $\varphi(y)$. The mathematical model describing this premise is based on the assumption that the Lipschitz condition is satisfied
\[
\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty.
\]
This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009}.
At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016}. 
An important property of global optimization problems is the fact that, in contrast to the problems of finding a local extremum, the global minimum is an integral characteristic of the problem being solved. Making sure that the point  $y^*\in D$ is a solution to the problem requires going beyond its neighborhood to the investigation of the entire search domain. As a result, when minimizing substantially multi-extremal functions, the numerical method must construct a coverage of the search domain. The number of nodes of this coverage increases exponentially with increasing dimensionality. 
This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solving.

>>>>

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
\usepackage{inputenc}
\usepackage[T2A,T1]{fontenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%

@@

There are several approaches to the construction of numerical methods for solving global optimization problems. 
Some algorithms are based on the idea of a multistart: launching a local search either from different starting points or with varying parameters. Local optimization methods have a high convergence rate. At the same time, one of the main problems in multistart schemes is the choice of starting points that would correspond to the regions of attraction of various local solutions. 
Machine learning methods can be successfully applied to solve this problem.  
For example, in \cite{RinnooyKan1987} methods of cluster analysis were used to select promising starting points.
In \cite{Cassioli2012} the area for starting the local method was allocated based on the classification of starting points using a support vector machine.

%Russian
Методы машинного обучения активно используются в сочетании с алгоритмами bayesian optimization, основанными на использовании probabilistic surrogate models of the objective function. Детальный обзор данного направления развития методов глобальной оптмизации представлен в \cite{Archetti2019,Zhigljavsky2021}.

@@

Note that the algorithms of the last class do not provide guaranteed convergence to the solution of the problem and are inferior to deterministic algorithms in terms of the quality of solution \cite{Kvasov2018,Sergeyev2018} (e.g., measured by the number of correctly solved problems from a particular set). Therefore, deterministic methods seem to be potentially more effective.

%Russian
Настоящая работа направлена на дальнейшее развитие эффективного детерминированного метода глобальной оптимизации, известного как \textit{the information-statistical global search algorithm} \cite{Strongin2000}. Указанная здесь книга содержит результаты теоретических исследований метода, имеющих непосредственное значение для его практической реализации. В частности, в ней детально рассмотрены вопросы сходимости, выбора параметров и условия остановки алгоритма, и т.д.
Отметим, что алгоритм глобального поиска первоначально был ориентирован на решение unconstrained optimization problems. Впоследствии он был обобщен для решения problems with non-convex constraints\cite{Barkalov2002} and multicriteria optimization problems \cite{Gergel2020}.
Одновременно с этим были предложены различные параллельные варианты указанных алгоритмов, которые могут быть использованы на современных суперкомпьютерах \cite{Barkalov2016,globalizerSystem,Strongin2018}.

@@

A typical situation for many applied problems is when a limited change in the vector of parameters $y$ causes a limited change in the values of $\varphi(y)$. The mathematical model describing this premise is based on the assumption that the Lipschitz condition is satisfied
\[
\left|\varphi(y')-\varphi(y'')\right|\leq L\left\|y'-y''\right\|,\; y',y'' \in D,\; 0<L<\infty.
\]
This assumption is typical for many approaches to the development of optimization algorithms \cite{Jones1993,Pinter1996,Zilinskas2008,Evtushenko2009,Sergeyev2020}.
At the same time, many known approaches are based on various methods of dividing the search domain into a system of subdomains and then choosing the most promising subdomain for placing the next trial (calculating the value of the objective function)  \cite{Jones2009,Zilinskas2010,Evtushenko2013,Kvasov2013,Paulavicius2016,Paulavicius2020}. 
An important property of global optimization problems is the fact that, in contrast to the problems of finding a local extremum, the global minimum is an integral characteristic of the problem being solved. Making sure that the point  $y^*\in D$ is a solution to the problem requires going beyond its neighborhood to the investigation of the entire search domain. As a result, when minimizing substantially multi-extremal functions, the numerical method must construct a coverage of the search domain. The number of nodes of this coverage increases exponentially with increasing dimensionality. 
This feature determines the high complexity of solving multiextremal optimization problems making dimensionality a critical factor that affects the complexity of their solving.
